\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs,array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}

\hypersetup{
  pdftitle={UOIF Reflection: IMO 2024--2025 with Official 2025 Results (Stepwise, Confidence)},
  pdfauthor={UOIF Working Note},
  pdfsubject={Unified Olympiad Information Framework (UOIF) Reflection},
  pdfkeywords={IMO, DeepMind, Evan Chen, AoPS, UOIF, reliability, canonical, scoring}
}

\newcommand{\Sx}{S(x)}
\newcommand{\Nx}{N(x)}
\newcommand{\Px}{\Psi(x)}
\newcommand{\post}{P(H\mid E,\beta)}
\newcommand{\pen}{\mathrm{Penalty}}
\newcommand{\conf}[2]{\textbf{Confidence:} #1\ (#2)}

\title{UOIF Reflection: IMO 2024--2025 with Official 2025 Results\\
Stepwise Recomputations and Confidence Accounting}
\author{UOIF Working Note}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Confirmed official IMO 2025 \emph{results} pages are live on \texttt{imo-official.org} (year info, country/individual results, statistics). We added canonical links for results primitives. Problems page remains pending. We recompute $\Px$ with updated parameters for 2025 results ($\alpha=0.12$, $\Nx=0.97$, $\beta=1.15$) yielding $\Px\approx 0.738$ (conservative), sufficient for promotion to \emph{Empirically Grounded} for results claims, while keeping problems \emph{Interpretive/Contextual} pending official problems/shortlist. 2024 remains \emph{Primitive/Empirically Grounded}.
\end{abstract}

\section{Canonical Links (2025 Results)}
\begin{itemize}[leftmargin=1.35em]
  \item Year info (medal thresholds): \url{https://imo-official.org/year_info.aspx?year=2025}
  \item Country results (team rankings/scores): \url{https://imo-official.org/year_country_r.aspx?year=2025}
  \item Individual results (participant scores/medals): \url{https://imo-official.org/year_individual_r.aspx?year=2025}
  \item Statistics: \url{https://imo-official.org/year_statistics.aspx?year=2025}
  \item Problems page: pending (no 2025 content found at time of writing).
\end{itemize}
Additional sources: DeepMind 2025 blog (certified gold) and solutions PDF; Evan Chen 2025 notes; AoPS 2025 threads (P1--P6); DeepMind 2024 solutions pages and blog; Evan site; historical mirror (context).

\section{Decision Model}
\[
\Px \;=\; \underbrace{\bigl[\alpha\,\Sx + (1-\alpha)\,\Nx\bigr]}_{\text{hybrid evidence}}
\cdot
\underbrace{\exp\!\Bigl(-\bigl[\lambda_1 R_{\text{authority}}+\lambda_2 R_{\text{verifiability}}\bigr]\Bigr)}_{\pen}
\cdot
\underbrace{\post}_{\text{posterior uplift}},
\quad \lambda_1=0.85,\ \lambda_2=0.15,\ \Sx=0.60.
\]

\section{IMO 2025 Results (Canonical live): Stepwise Recalculation}
\subsection*{Step 1: Sources and roles}
Canonical results pages (above) + DeepMind 2025 (certified) + Evan 2025 + AoPS (context for problems).\\
\conf{0.98}{Official results pages live; multi-source triangulation.}

\subsection*{Step 2: External reliability and allocation}
Set $\Nx=0.97$ (official uplift), $\alpha=0.12$ (favor canonical). Hybrid:
\[
O_{\text{hybrid}} = 0.12\cdot 0.60 + 0.88\cdot 0.97 = 0.9256.
\]
\conf{0.96}{Authority and verifiability high for results; allocation reflects canonical priority.}

\subsection*{Step 3: Penalty}
Results canonical live; problems pending. We retain a cautious penalty for mixed state:
\[
\pen=\exp\!\bigl(-[0.85\cdot 0.15 + 0.15\cdot 0.05]\bigr)=\exp(-0.135)\approx 0.8737.
\]
\conf{0.85}{Conservative carry-over while problems remain non-canonical.}

\subsection*{Step 4: Posterior uplift}
For results with canonical uplift: $P(H\mid E)=0.85$--$0.90$, $\beta=1.15$. We adopt a conservative calibrated value
\[
\post \approx 0.913\quad(\text{within }[0.85,1.0]).
\]
\conf{0.85}{Conservative setting to avoid over-promotion prior to full problems/shortlist.}

\subsection*{Step 5: Computation and label}
Conservative (operational):
\[
\Px \approx 0.9256 \times 0.8737 \times 0.913 \approx 0.738.
\]
Nominal (for reference; $\post=0.945$):
\[
\Px \approx 0.9256 \times 0.8737 \times 0.945 \approx 0.764.
\]
\textbf{Label:} \emph{Empirically Grounded} for 2025 results primitives.\\
\conf{0.90}{Both conservative and nominal exceed 0.70 threshold; we report the conservative value.}

\subsection*{Step 6: Confidence summary}
\begin{center}
\renewcommand{\arraystretch}{1.12}
\begin{tabular}{@{}lcl@{}}
\toprule
Step & Value/Decision & Confidence (rationale)\\
\midrule
Sources/roles & Canonical results + expert + community & 0.98 (official live)\\
$N(x)$, $\alpha$ & $N=0.97$, $\alpha=0.12$ & 0.96 (canonical uplift)\\
Penalty & $0.8737$ & 0.85 (mixed canonical state)\\
Posterior & $0.913$ (conservative), $0.945$ (nominal) & 0.85 (policy-consistent)\\
$\Px$ & $0.738$ (conservative), $0.764$ (nominal) & 0.90 (robust above threshold)\\
\bottomrule
\end{tabular}
\end{center}

\section{IMO 2025 Problems (Pending canonical): Stepwise Recalculation}
\subsection*{Step 1: Sources and roles}
Evan 2025 PDF + AoPS per-problem; DeepMind 2025 (expert interpretive). Canonical problems/shortlist pending.\\
\conf{0.88}{Strong expert corroboration; non-canonical primitives.}

\subsection*{Step 2: External reliability and allocation}
Set $\Nx\in[0.88,0.90]$, $\alpha\in[0.15,0.20]$ (favor interpretive). Example midpoint: $\Nx=0.89$, $\alpha=0.17$ gives
\[
O_{\text{hybrid}}=0.17\cdot 0.60 + 0.83\cdot 0.89=0.8417.
\]
\conf{0.85}{Pre-canonical envelope for problems.}

\subsection*{Step 3: Penalty and posterior}
Problems not canonical: $\pen=\exp(-[0.85\cdot 0.25+0.15\cdot 0.10])\approx 0.7965$ (stricter). Posterior uplift modest: $\post=0.90\times 1.05=0.945$.\\
\conf{0.80}{Higher authority penalty; modest uplift.}

\subsection*{Step 4: Computation and label}
Illustrative midpoint:
\[
\Px \approx 0.8417 \times 0.7965 \times 0.945 \approx 0.633.
\]
Range over $\alpha\in[0.15,0.20],\ \Nx\in[0.88,0.90]$: $\Px\approx 0.60$--$0.65$.\\
\textbf{Label:} \emph{Interpretive/Contextual; strong corroboration; pending official}.\\
\conf{0.85}{Range stable; promotion expected on canonical posting.}

\section{IMO 2024 (Enhanced; P1/P2/P4): Stepwise Recalculation}
\subsection*{Step 1: Sources and roles}
DeepMind 2024 blog + solutions pages (P1/P2/P4), Evan site, established archives/mirrors; canonical availability varies.\\
\conf{0.90}{Multiple expert artifacts; stable mirrors.}

\subsection*{Step 2: External reliability and allocation}
$\Nx=0.96$, $\alpha\in[0.10,0.15]$. Hybrid $O_{\text{hybrid}}=0.96-0.36\alpha$.\\
\conf{0.88}{Established context and coverage.}

\subsection*{Step 3: Penalty and posterior}
$\pen=\exp(-[0.85\cdot 0.10+0.15\cdot 0.05])\approx 0.9117$, $\post=0.90\times 1.05=0.945$.\\
\conf{0.85}{Lower authority penalty vs 2025 problems.}

\subsection*{Step 4: Computation and label}
\[
\alpha=0.10:\ \Px \approx 0.9240\times 0.9117\times 0.945 \approx 0.796,\quad
\alpha=0.15:\ \Px \approx 0.9060\times 0.9117\times 0.945 \approx 0.781.
\]
\textbf{Label:} \emph{Primitive/Empirically Grounded}.\\
\conf{0.88}{Range consistent with artifacts in hand.}

\section{Summary Table}
\begin{center}
\renewcommand{\arraystretch}{1.12}
\begin{tabular}{@{}lcccccc@{}}
\toprule
Scope & $\alpha$ & $\Nx$ & \pen & \post & $\Px$ & Label\\
\midrule
2025 Results (cons.) & 0.12 & 0.97 & 0.8737 & 0.913 & 0.738 & Empirically Grounded\\
2025 Results (nom.) & 0.12 & 0.97 & 0.8737 & 0.945 & 0.764 & (ref.)\\
2025 Problems & 0.15--0.20 & 0.88--0.90 & 0.7965 & 0.945 & 0.60--0.65 & Interpretive/Contextual\\
2024 P1/P2/P4 & 0.10--0.15 & 0.96 & 0.9117 & 0.945 & 0.781--0.796 & Primitive/Empirically Grounded\\
\bottomrule
\end{tabular}
\end{center}

\section*{Changelog}
Added canonical 2025 results links (year info, country, individual, statistics). Recomputed 2025 results with $\alpha=0.12$, $\Nx=0.97$, $\beta=1.15$; reported conservative $\Px\approx 0.738$ (promotion achieved for results). 2025 problems remain pending; 2024 unchanged.

\end{document}