% uoif_reflection.tex
\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs,array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}

\hypersetup{
  pdftitle={UOIF Reflection: IMO 2024--2025 with Official 2025 Results},
  pdfauthor={UOIF Working Note},
  pdfsubject={UOIF Reflection: recomputations and confidence},
  pdfkeywords={IMO, DeepMind, Evan Chen, AoPS, UOIF, canonical, scoring}
}

\newcommand{\Sx}{S(x)}
\newcommand{\Nx}{N(x)}
\newcommand{\Px}{\Psi(x)}
\newcommand{\post}{P(H\mid E,\beta)}
\newcommand{\pen}{\mathrm{Penalty}}
\newcommand{\conf}[2]{\textbf{Confidence:} #1\ (#2)}

\title{UOIF Reflection: IMO 2024--2025 with Official 2025 Results\\
Stepwise Recomputations and Confidence Accounting}
\author{UOIF Working Note}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Official IMO 2025 \emph{results} pages (year info, country/individual results, statistics) are live on \texttt{imo-official.org}; we integrated these canonical links and recomputed $\Px$ for 2025 results using $\alpha=0.12$, $\Nx=0.97$, $\beta=1.15$. Conservative $\Px\approx 0.738$ (nominal $\approx 0.764$) clears the threshold for \emph{Empirically Grounded} results claims. 2025 problems remain \emph{Interpretive/Contextual} pending official problems/shortlist. 2024 stays \emph{Primitive/Empirically Grounded}.
\end{abstract}

\section{Decision Model}
\[
\Px = \underbrace{[\alpha\,\Sx + (1-\alpha)\,\Nx]}_{\text{hybrid}}
\cdot
\underbrace{\exp(-[\lambda_1 R_{\text{authority}}+\lambda_2 R_{\text{verifiability}}])}_{\pen}
\cdot
\underbrace{\post}_{\text{posterior}},
\quad \lambda_1=0.85,\ \lambda_2=0.15,\ \Sx=0.60.
\]

\section{IMO 2025 Results (Canonical live): Stepwise}
\subsection*{Step 1: Sources/roles}
Canonical results:
\begin{itemize}[leftmargin=1.35em]
  \item Year info: \url{https://imo-official.org/year_info.aspx?year=2025}
  \item Country results: \url{https://imo-official.org/year_country_r.aspx?year=2025}
  \item Individual results: \url{https://imo-official.org/year_individual_r.aspx?year=2025}
  \item Statistics: \url{https://imo-official.org/year_statistics.aspx?year=2025}
\end{itemize}
Expert interpretive: DeepMind 2025 blog \& solutions PDF; Evan 2025; AoPS (problems).\\
\conf{0.98}{Official results live; triangulated with expert/community.}

\subsection*{Step 2: Reliability/Allocation}
Set $\Nx=0.97$ (canonical uplift), $\alpha=0.12$.
\[
O_{\text{hybrid}}=0.12\cdot 0.60 + 0.88\cdot 0.97 = 0.9256.
\]
\conf{0.96}{High authority/verifiability for results; canonical prioritized.}

\subsection*{Step 3: Penalty}
Mixed state (results canonical; problems pending): keep conservative
\[
\pen=\exp(-[0.85\cdot 0.15 + 0.15\cdot 0.05])=\exp(-0.135)\approx 0.8737.
\]
\conf{0.85}{Conservative until problems/shortlist post.}

\subsection*{Step 4: Posterior}
$P(H\mid E)=0.85$--$0.90$, $\beta=1.15$. Conservative calibration:
\[
\post \approx 0.913\quad(\text{nominal }0.945).
\]
\conf{0.85}{Avoids over-promotion pre-full-canonical.}

\subsection*{Step 5: Compute \& Label}
Conservative:
\[
\Px \approx 0.9256 \times 0.8737 \times 0.913 \approx 0.738.
\]
Nominal:
\[
\Px \approx 0.9256 \times 0.8737 \times 0.945 \approx 0.764.
\]
\textbf{Label:} \emph{Empirically Grounded} for 2025 results primitives.\\
\conf{0.90}{Both values exceed 0.70; conservative reported operationally.}

\section{IMO 2025 Problems (Pending canonical): Stepwise}
\subsection*{Step 1: Sources/roles}
Evan 2025 PDF + AoPS per-problem; DeepMind 2025 (expert interpretive). Problems/shortlist pending.\\
\conf{0.88}{Strong expert corroboration; non-canonical primitives.}

\subsection*{Step 2--4: Reliability/Allocation/Penalty/Posterior}
$\Nx\in[0.88,0.90]$, $\alpha\in[0.15,0.20]$; example midpoint $\Nx=0.89$, $\alpha=0.17$:
\[
O_{\text{hybrid}}=0.17\cdot 0.60 + 0.83\cdot 0.89=0.8417.
\]
Stricter penalty for non-canonical problems:
\[
\pen=\exp(-[0.85\cdot 0.25+0.15\cdot 0.10])\approx 0.7965,\quad \post=0.945.
\]
\conf{0.80}{Higher authority penalty; modest uplift.}

\subsection*{Step 5: Compute \& Label}
Midpoint:
\[
\Px \approx 0.8417 \times 0.7965 \times 0.945 \approx 0.633.
\]
Range: $\Px\approx 0.60$--$0.65$ over $\alpha\in[0.15,0.20]$, $\Nx\in[0.88,0.90]$.\\
\textbf{Label:} \emph{Interpretive/Contextual; strong corroboration; pending official}.\\
\conf{0.85}{Stable range; promotion expected when canonical posts.}

\section{IMO 2024 (Enhanced; P1/P2/P4): Stepwise}
\subsection*{Step 1: Sources/roles}
DeepMind 2024 blog + solutions pages (P1/P2/P4), Evan site, established archives/mirrors.\\
\conf{0.90}{Multiple expert artifacts; stable mirrors.}

\subsection*{Step 2--4: Reliability/Allocation/Penalty/Posterior}
$\Nx=0.96$, $\alpha\in[0.10,0.15]$, $\pen=\exp(-[0.85\cdot 0.10+0.15\cdot 0.05])\approx 0.9117$, $\post=0.945$.\\
\conf{0.85}{Lower authority penalty vs 2025 problems.}

\subsection*{Step 5: Compute \& Label}
\[
\alpha=0.10:\ \Px \approx 0.9240\times 0.9117\times 0.945 \approx 0.796,\quad
\alpha=0.15:\ \Px \approx 0.9060\times 0.9117\times 0.945 \approx 0.781.
\]
\textbf{Label:} \emph{Primitive/Empirically Grounded}.\\
\conf{0.88}{Consistent with artifacts in hand.}

\section{Verification Methodology (brief)}
We treat classical solvers and official artifacts as baselines:
\begin{enumerate}[leftmargin=1.35em]
  \item Train/curate interpretive artifacts (expert notes, formal solutions).
  \item Use canonical pages (official results/problems/shortlist) as ground truth for primitives.
  \item Compare/align interpretive claims; quantify error/consistency and promotion readiness.
\end{enumerate}

\section{Summary Table}
\begin{center}
\renewcommand{\arraystretch}{1.12}
\begin{tabular}{@{}lcccccc@{}}
\toprule
Scope & $\alpha$ & $\Nx$ & \pen & \post & $\Px$ & Label\\
\midrule
2025 Results (cons.) & 0.12 & 0.97 & 0.8737 & 0.913 & 0.738 & Empirically Grounded\\
2025 Results (nom.)  & 0.12 & 0.97 & 0.8737 & 0.945 & 0.764 & (reference)\\
2025 Problems        & 0.15--0.20 & 0.88--0.90 & 0.7965 & 0.945 & 0.60--0.65 & Interpretive/Contextual\\
2024 P1/P2/P4        & 0.10--0.15 & 0.96 & 0.9117 & 0.945 & 0.781--0.796 & Primitive/Empirically Grounded\\
\bottomrule
\end{tabular}
\end{center}

\section*{Changelog}
Added canonical 2025 results links; recomputed 2025 results with $\alpha=0.12$, $\Nx=0.97$, $\beta=1.15$; reported conservative $\Px\approx 0.738$; promoted results to \emph{Empirically Grounded}. 2025 problems pending; 2024 unchanged.

\end{document}